Computing Entropies With Nested Sampling

### 2017-06-27 paper/ms.tex
Couple of lines

NS evolves a population of
$N$ particles in the parameter space.
The particles are initially initialised from the prior $\pi(\x)$, and the
particle with the lowest likelihood, $L^*_1$, is found and its details are
recorded as output. This worst particle is then discarded
and replaced by a new particle
drawn from the prior $\pi$ but subject to the constraint that its likelihood
must be above $L^*_1$ (this is usually achieved by cloning a surviving particle
and evolving it with MCMC). This process is repeated, resulting in an
increasing sequence of likelihoods
\begin{align}
L^*_1, L^*_2, L^*_3..., 
\end{align}

@@

This example demonstrates how the algorithm may be used to determine the
entropy for a quantity whose distribution can only be sampled.
Consider the basic statistics problem of inferring a quantity $\mu$ from
100 observations $\d = \{d_1, ..., d_{100}\}$ whose
probability distribution (conditional on $\mu$) is
\begin{align}
p(d_i | \mu) &\sim \textnormal{Normal}(\mu, 1).
\end{align}
If the prior for $\mu$ is Normal$(0, 10^2)$, then the posterior is
\begin{align}
\mu | \d &\sim \textnormal{Normal}\left(
                                       \frac{1}{100}\sum_{i=1}^{100} d_i,
                                       \left[\frac{10}{\sqrt{10001}}\right]^2
                                       \right).
\end{align}
The true value of the differential entropy $H(\d)$ is available analytically
in this case, and its value is $146.499$ nats. However, in most situations, the
entropy of a marginal distribution is not available in closed form.

@@

As the algorithm lets us calculate the entropy of any distribution which can
be sampled by MCMC, there are several options. The mutual information
can be written in the following three ways
\begin{align}
I(x; y) &= H(x) + H(y) - H(x, y) \\
I(x; y) &= H(y) - H(y|x) \\
I(x; y) &= H(x) - H(x|y) \\
I(x; y) &= \int D_{\rm KL}\left(p(x|y) || p(x)\right) \, dy.
\end{align}
In a Bayesian problem where $x$ is an unknown parameter and $y$ is data, the
first and second formulas would be very costly to compute, because of the
high dimensional probability distribution for the dataset.

@@

\acknowledgments{Acknowledgements}
It is a pleasure to thank the following people for interesting and helpful
conversations about this topic: Ruth Angus (Flatiron Institute),
Ewan Cameron (Oxford), James Curran (Auckland),
David Hogg (NYU), Kevin Knuth (SUNY Albany),
Thomas Lumley (Auckland),
Iain Murray (Edinburgh), Jared Tobin (jtobin.io).
This work was supported by Centre for eResearch
at the University of Auckland.

>>>>

NS evolves a population of
$N$ particles in the parameter space.
The particles are initially initialised from the prior $\pi(\x)$, and the
particle with the lowest likelihood, $L^*_1$, is found and its details are
recorded as output. This worst particle is then discarded
and replaced by a new particle
drawn from the prior $\pi$ but subject to the constraint that its likelihood
must be above $L^*_1$ (this is usually achieved by cloning a surviving particle
and evolving it with MCMC). This process is repeated, resulting in an
increasing sequence of likelihoods
\begin{align}
L^*_1, L^*_2, L^*_3,..., 
\end{align}

@@

This example demonstrates how the algorithm may be used to determine the
entropy for a quantity whose distribution can only be sampled, in this case,
a prior distribution over datasets.

Consider the basic statistics problem of inferring a quantity $\mu$ from
100 observations $\d = \{d_1, ..., d_{100}\}$ whose
probability distribution (conditional on $\mu$) is
\begin{align}
p(d_i | \mu) &\sim \textnormal{Normal}(\mu, 1).
\end{align}
If the prior for $\mu$ is Normal$(0, 10^2)$, then the posterior is
\begin{align}
\mu | \d &\sim \textnormal{Normal}\left(
                                       \frac{1}{100}\sum_{i=1}^{100} d_i,
                                       \left[\frac{10}{\sqrt{10001}}\right]^2
                                       \right).
\end{align}
The true value of the differential entropy $H(\d)$ is available analytically
in this case, and its value is $146.499$ nats. However, in most situations, the
entropy of a marginal distribution is not available in closed form.

@@

As the algorithm lets us calculate the entropy of any distribution which can
be sampled by MCMC, there are several options. The mutual information
can be written in the following three ways
\begin{align}
I(x; y) &= H(x) + H(y) - H(x, y) \\
I(x; y) &= H(y) - H(y|x) \\
I(x; y) &= H(x) - H(x|y) \\
I(x; y) &= \int p(y) \, D_{\rm KL}\left(p(x|y) \,||\, p(x)\right) \, dy.
\end{align}
In a Bayesian problem where $x$ is an unknown parameter and $y$ is data, the
first and second formulas would be very costly to compute, because of the
high dimensional probability distribution for the dataset.

@@

\acknowledgments{Acknowledgements}
It is a pleasure to thank the following people for interesting and helpful
conversations about this topic: Ruth Angus (Flatiron Institute),
Ewan Cameron (Oxford), James Curran (Auckland),
David Hogg (NYU), Kevin Knuth (SUNY Albany),
Thomas Lumley (Auckland),
Iain Murray (Edinburgh), Jared Tobin ({\tt jtobin.io}).
This work was supported by Centre for eResearch
at the University of Auckland.
